---
title: "p8105_hw3_zf2212"
author: "Catherine"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)
library(tidyverse)
library(p8105.datasets)
```

## Problem 1 BRFSS

First, do some data cleaning:

```{r import and tidy data}
# import
data("brfss_smart2010")
# clean
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% # use appropriate variable names
  filter(topic == "Overall Health") %>% # filter topic 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), !is.na("response")) %>% 
  # filter response from excellent to poor
  mutate(response, response = as.factor(response)) %>% 
  arrange(response) %>%  # mutate response as factor and arrange order
  select(year, locationabbr, locationdesc, response, data_value, everything())
```

__1.1__

In 2002, which states were observed at 7 locations?

```{r observed 7}
# filter the 2002 data
brfss2002 = brfss %>% 
  filter(year == 2002)
# summarize the number of locations
n_obs2002 = brfss2002 %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) 
# find the state observed 7 location
obs7 = n_obs2002 %>% 
  filter(n_location == 7) 
obs7[1]
```

There are three states were observed at 7 locations, including __CT, FL and NC.__

__1.2__

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r spaghetti plot}
# arrange the data
brfss_nloc = brfss %>%
  group_by(year, locationabbr) %>% 
  summarize(n_loc = n_distinct(locationdesc))
# create the plot
brfss_nloc %>% 
  ggplot(aes(x = year, y = n_loc, color = locationabbr )) + 
  geom_point(size = 0.8) +
  geom_line(size = 0.5) +
  labs(
    title = "2002-2010 Number of Locations Observed in Each States",
    x = "Year",
    y = "Number of Locations",
    caption = "Data from the data.gov") + 
  scale_color_hue(name = "States") +
  theme_bw() +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), legend.text = element_text(size = 7)) +
  guides(col = guide_legend(nrow = 3))
```

__1.3__

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r table for ny} 
# filter ny data
brfss_ny = brfss %>% 
  filter(year == 2002 | year == 2006 | year == 2010, locationabbr == "NY", response == "Excellent")  %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  spread(key = "response", value = "data_value") %>% 
  group_by(year) %>% # group data by year
  janitor::clean_names()  %>% 
  summarize(excellent_mean = mean(excellent, na.rm = TRUE), 
            excellent_sd = sd(excellent, na.rm = TRUE)) # summarize mean and sd
brfss_ny 
```

__1.4__

For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time. 
(five-panel plot -  facet_grid() )
```{r five panel plot}
# arrange the data 
brfss_average = brfss %>% 
  group_by(year, locationabbr) %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  spread(key = "response", value = "data_value") %>% 
  janitor::clean_names() %>% # compute the average proportion
  summarize(excellent_average = mean(excellent, na.rm = TRUE),
            fair_average = mean(fair, na.rm = TRUE),
            good_average = mean(good, na.rm = TRUE),
            poor_average = mean(poor, na.rm = TRUE),
            very_good_ave = mean(very_good, na.rm = TRUE)) 
brfss_average

# arrange data for the plot
average_plot = 
  brfss_average %>% 
  group_by(year, locationabbr) %>% 
  gather(key = "response", value = "propotion", excellent_average:very_good_ave)

# set up a lookup table for the plot
fivepanel = c(excellent_average = "Excellent",
              fair_average = "Fair",
              good_average = "Good",
              poor_average = "Poor",
              very_good_ave = "Very Good")

# make fivepanel plot
average_plot %>% 
ggplot(aes(x = year , y = propotion, color = locationabbr)) +
  geom_point(size = 0.5) + 
  geom_line(size = 0.3) +
  facet_grid(~response, labeller = labeller(response = fivepanel)) +
  labs(
    title = "2002-2010 Distribution of Response Proportion Mean at State-level",
    x = "Year",
    y = "Response Proportiaon Mean",
    caption = "Data from data.gov") + 
  scale_color_hue(name = "States") +
  theme_bw() +
  theme(legend.position = "bottom", plot.title = element_text(hjust=0.5), legend.text = element_text(size = 7)) +
  guides(col = guide_legend(nrow = 3))
```


## Problem 2 Instacart

__2.1__

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

```{r describe data}
# import and clean
data("instacart")
instacart %>% 
  janitor::clean_names()
# decribe data
names(instacart) # variables name
dim(instacart) # size
head(instacart, 5) 
tail(instacart, 5)
skimr::skim(instacart) # look at the data
instacart %>% # find unique users numbers
  distinct(user_id) %>% 
  nrow()
```

__Description__

The dataset originally comes from _instacart.com_. The version we used for this analysis is cleaned version from p8105.datasets. There are `r dim(instacart)` (observations, variables) in the dataset. Each row in the dataset is a product from an order and a single order per user. There are 131,209 unique users. 

The variables include information of `r names(instacart)`. There are no missing value in each column. 

For the following analysis, the variables of interest includes??? (order_id, product_id, reordered, user_id, order_number) as they are some key information for us to figure out the shopping behaviors of the users in _Instacart_. 

In __reordered__, 1 represent this prodcut has been ordered by this user in the past and 0 represents it was not ordered before. For example, for _order#1_, four out of eight products were reordered products. 

In **order_dow**, the numbers in the column show the day of the week on which the order was placed. As information about the start day of the week is not shown, we assume the first day of the week is Sunday, which means 0 represents Sunday. 

__2.2__ (comment on each result)

How many aisles are there, and which aisles are the most items ordered from?

```{r aisles count}
# make table of aisle 
aisles_data = instacart %>% 
  select(order_id, product_id, user_id, aisle_id, aisle, department) %>% 
  group_by(aisle, aisle_id, department) %>% 
  summarize(items_n = n()) %>% 
  arrange(desc(items_n)) 
# count the number of aisles
n_aisles = nrow(aisles_data)
# find aisles that most items ordered from
aisles_data[1,1]
```

There are __`r n_aisles`__ aisles and __`r aisles_data[1,1]`__ is the aisle which the most items ordered from.

__2.3__

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it. 

```{r aisles plot}
# arrange data for the plot
aisles_plot = instacart %>% 
  select(aisle_id, aisle) %>% 
  group_by(aisle) %>% 
  summarize(items_n = n()) %>% 
  mutate(items_n, items_n = items_n/1000) 
# make the plot
aisles_plot %>% 
  ggplot(aes(x = aisle, y = items_n, color = aisle)) + # set the axis
  geom_point(aes(size = items_n)) + 
  labs(title = "The Number of Items in Each Aisle",
       x = "Aisle",
       y = "Number of Items (x1000)",
       caption = "The size of point reflects the number of items in each aisle.") +
  #scale_x_discrete(expand = expand_scale(add = 4)) +
  theme_bw() +
  theme(text = element_text(size = 10), plot.title = element_text(hjust = 0.5), legend.position = "hide", axis.text.x = element_text(angle = 90, hjust = 1)) 
```

As the number of each item range from 287 to 150609, the number was divided by 1000 in order to make the plot looks clear. Also, differentiate the number by the size of point make it more easy to notice the aisle with large nubmer. 

__2.4__

Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. -- one table, showing one item from each of these three aisles.

```{r popular item}
# count the buying times of each product
popitem = instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(times = n()) 
# fine most popular in each aisle
baking = popitem %>% 
  filter(aisle == "baking ingredients") %>% 
  arrange(desc(times)) 
  
dfc = popitem %>% 
  filter(aisle == "dog food care") %>% 
  arrange(desc(times)) 

pkvege = popitem %>% 
  filter(aisle == "packaged vegetables fruits") %>% 
  arrange(desc(times))

# create table
rbind(baking[1,], dfc[1,], pkvege[1,])
```
The most popular item in baking ingredients is __Light Brown Sugar__. 

The most popular item in dog food care is __Snack Sticks Chicken & Rice Recipe Dog Treats__.

The most popular item in packaged vegetables fruits is __Organic Baby Spinach__.

__2.5__
Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).


## Problem 3 NY NOAA
This problem uses the NY NOAA data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package (it’s called ny_noaa).

```{r}
data(ny_noaa)

```

The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.  (obtain proportions of the missing values by column and comment about the degree to which the data is missing, maybe connect to the background of the data)
Then, do or answer the following (commenting on the results of each):

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); (using geom_hex, we see how densities of tmin/tmax combinations; when using geom_smooth, we see more of the relationship between tmin and tmax. / error geom_hex:downloaded the package hexbin)and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
