---
title: "p8105_hw3_zf2212"
author: "Catherine"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
library(tidyverse)
library(p8105.datasets)
library(patchwork)
```

## Problem 1 BRFSS

First, do some data cleaning:

```{r import and tidy data}
# import
data("brfss_smart2010")
# clean
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% # use appropriate variable names
  filter(topic == "Overall Health") %>% # filter topic 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), !is.na("response")) %>% 
  # filter response from excellent to poor
  mutate(response, response = as.factor(response)) %>% 
  arrange(response) %>%  # mutate response as factor and arrange order
  select(year, locationabbr, locationdesc, response, data_value, everything())
```

### 1.1

In 2002, which states were observed at 7 locations?

```{r observed 7}
# filter the 2002 data
brfss2002 = brfss %>% 
  filter(year == 2002)
# summarize the number of locations
n_obs2002 = brfss2002 %>% 
  group_by(locationabbr) %>% 
  summarize(n_location = n_distinct(locationdesc)) 
# find the state observed 7 location
obs7 = n_obs2002 %>% 
  filter(n_location == 7) 
obs7[1]
```

There are three states were observed at 7 locations, including __CT, FL and NC.__

### 1.2

Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.

```{r spaghetti plot, fig.width = 10,fig.height = 11}
# arrange the data
brfss_nloc = brfss %>%
  group_by(year, locationabbr) %>% 
  summarize(n_loc = n_distinct(locationdesc))
# create the plot
brfss_nloc %>% 
  ggplot(aes(x = year, y = n_loc, color = locationabbr )) + 
  geom_point(size = 0.8) +
  geom_line(size = 0.5) +
  labs(
    title = "2002-2010 Number of Locations Observed in Each States",
    x = "Year",
    y = "Number of Locations",
    caption = "Data from the data.gov") + 
  scale_color_hue(name = "States") +
  theme_bw() +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), legend.text = element_text(size = 7)) +
  guides(col = guide_legend(nrow = 3))
```

### 1.3

Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.

```{r table for ny} 
# filter ny data
brfss_ny = brfss %>% 
  filter(year == 2002 | year == 2006 | year == 2010, locationabbr == "NY", response == "Excellent")  %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  spread(key = "response", value = "data_value") %>% 
  group_by(year) %>% # group data by year
  janitor::clean_names()  %>% 
  summarize(excellent_mean = mean(excellent, na.rm = TRUE), 
            excellent_sd = sd(excellent, na.rm = TRUE)) # summarize mean and sd
brfss_ny 
```

### 1.4

For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time. 
(five-panel plot -  facet_grid() )
```{r five panel plot, fig.width = 10,fig.height = 11}
# arrange the data 
brfss_average = brfss %>% 
  group_by(year, locationabbr) %>% 
  select(year, locationabbr, locationdesc, response, data_value) %>% 
  spread(key = "response", value = "data_value") %>% 
  janitor::clean_names() %>% # compute the average proportion
  summarize(excellent_average = mean(excellent, na.rm = TRUE),
            fair_average = mean(fair, na.rm = TRUE),
            good_average = mean(good, na.rm = TRUE),
            poor_average = mean(poor, na.rm = TRUE),
            very_good_ave = mean(very_good, na.rm = TRUE)) 
brfss_average

# arrange data for the plot
average_plot = 
  brfss_average %>% 
  group_by(year, locationabbr) %>% 
  gather(key = "response", value = "propotion", excellent_average:very_good_ave)

# set up a lookup table for the plot
fivepanel = c(excellent_average = "Excellent",
              fair_average = "Fair",
              good_average = "Good",
              poor_average = "Poor",
              very_good_ave = "Very Good")

# make fivepanel plot
average_plot %>% 
ggplot(aes(x = year , y = propotion, color = locationabbr)) +
  geom_point(size = 0.5) + 
  geom_line(size = 0.3) +
  facet_grid(~response, labeller = labeller(response = fivepanel)) +
  labs(
    title = "2002-2010 Distribution of Response Proportion Mean at State-level",
    x = "Year",
    y = "Response Proportiaon Mean",
    caption = "Data from data.gov") + 
  scale_color_hue(name = "States") +
  theme_bw() +
  theme(legend.position = "bottom", plot.title = element_text(hjust=0.5), legend.text = element_text(size = 7)) +
  guides(col = guide_legend(nrow = 3))
```


## Problem 2 Instacart

### 2.1

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illstrative examples of observations.

```{r describe data}
# import and clean
data("instacart")
instacart %>% 
  janitor::clean_names()
# decribe data
names(instacart) # variables name
dim(instacart) # size
head(instacart, 5) 
tail(instacart, 5)
skimr::skim(instacart) # look at the data
instacart %>% # find unique users numbers
  distinct(user_id) %>% 
  nrow()
```

__Description__

The dataset originally comes from _instacart.com_. The version we used for this analysis is cleaned version from p8105.datasets. There are `r dim(instacart)` (observations, variables) in the dataset. Each row in the dataset is a product from an order and a single order per user. There are 131,209 unique users. 

The variables include information of `r names(instacart)`. There are no missing value in each column. 

For the following analysis, the variables of interest includes __aisle, product_name and department__ as they are some key information for us to figure out the shopping behaviors of the users in _Instacart_. 

In __reordered__, 1 represent this prodcut has been ordered by this user in the past and 0 represents it was not ordered before. For example, for _order#1_, four out of eight products were reordered products. 

In **order_dow**, the numbers in the column show the day of the week on which the order was placed. As information about the start day of the week is not shown, we assume the first day of the week is Sunday, which means 0 represents Sunday. 

### 2.2

How many aisles are there, and which aisles are the most items ordered from?

```{r aisles count}
# make table of aisle 
aisles_data = instacart %>% 
  select(order_id, product_id, user_id, aisle_id, aisle, department) %>% 
  group_by(aisle, aisle_id, department) %>% 
  summarize(items_n = n()) %>% 
  arrange(desc(items_n)) 
# count the number of aisles
n_aisles = nrow(aisles_data)
# find aisles that most items ordered from
aisles_data[1,1]
```

There are __`r n_aisles`__ aisles and __`r aisles_data[1,1]`__ is the aisle which the most items ordered from.

### 2.3

Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it. 

```{r aisles plot, fig.width = 20,fig.height = 20}
# arrange data for the plot
aisles_plot = instacart %>% 
  select(aisle_id, aisle) %>% 
  group_by(aisle) %>% 
  summarize(items_n = n())
  
# make the plot
aisles_plot %>% 
  ggplot(aes(x = reorder(aisle, -items_n), y = items_n, fill = aisle)) + # set the axis
  geom_bar(stat = "identity") + 
  labs(title = "The Number of Items in Each Aisle",
       x = "Aisle",
       y = "Number of Items",
       caption = "Data from instacart.com") + 
  coord_flip() + # flip x and y
  scale_y_continuous(breaks = c(0,5000,10000,75000,100000,150000), expand = c(0, 1000)) + # reset the breaks
  theme_bw() +
  theme(text = element_text(size = 6), plot.title = element_text(hjust = 0.5), legend.position = "hide", title = element_text(size = 10, face = 'bold')) # set theme
```

As the number of each item range from 287 to 150609, which is very big, I break the Number of Items manually to make it clear. Also, differentiate each aisle by color. Instead of keeping the legend, I hided it and flip the x and y axis, which creats more space for the name of each aisle. It is easier to look compared to check the legend. By reordering the bars, max and min value are easy to read.

### 2.4

Make a table showing the most popular item in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. -- one table, showing one item from each of these three aisles.

```{r popular item}
# count the buying times of each product
popitem = instacart %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>% 
  summarize(times = n()) 
# fine most popular in each aisle
baking = popitem %>% 
  filter(aisle == "baking ingredients") %>% 
  arrange(desc(times)) 
  
dfc = popitem %>% 
  filter(aisle == "dog food care") %>% 
  arrange(desc(times)) 

pkvege = popitem %>% 
  filter(aisle == "packaged vegetables fruits") %>% 
  arrange(desc(times))

# create table
rbind(baking[1,], dfc[1,], pkvege[1,])
```
The most popular item in baking ingredients is __Light Brown Sugar__. 

The most popular item in dog food care is __Snack Sticks Chicken & Rice Recipe Dog Treats__.

The most popular item in packaged vegetables fruits is __Organic Baby Spinach__.

### 2.5

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. 

```{r apple and ice cream}
placic = instacart %>% 
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  select(product_name, order_dow, order_hour_of_day) %>% 
  summarize(hr_mean = round(mean(order_hour_of_day), digits = 3)) %>% 
  spread(key = order_dow, value = hr_mean)
placic
```
The table shows the mean hour of the day Pink Lady Apples and Coffee Ice Cream are ordered on each day of week. We can see that __Ice Cream were most ordered later than Apples in most of the days__. However, as mean is greatly affected by extreme value, we need more analysis to draw the conclusion.

## Problem 3 NY NOAA
### 3.1

Write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue.

```{r noaa description}
data(ny_noaa)
ny_noaa %>% 
  janitor::clean_names()
# decribe data
names(ny_noaa) # variables name
dim(ny_noaa) # size
dis_id = nrow(distinct(ny_noaa, id)) # distinct id
head(ny_noaa, 5) 
tail(ny_noaa, 5)
skimr::skim(ny_noaa) # look at the data
# obtain the proportion of missing value
tb_miss = ny_noaa %>%
  summarize(prcp_na = sum(is.na(prcp))/n(),
            snow_na = sum(is.na(snow))/n(),
            snwd_na = sum(is.na(snwd))/n(),
            tmax_na = sum(is.na(tmax))/n(),
            tmin_na = sum(is.na(tmin))/n())
tb_miss
```

The dataset originally comes from _NOAA National Climatic Data Center_. The version we used for this analysis is cleaned version from p8105.datasets. There are __`r dim(ny_noaa)` (observations, variables)__ in the dataset. It includes five core variables for all New York state weather stations from January 1, 1981 through December 31, 2010. There are `r dis_id` distinct Weather stations. 

The variables include information of __`r names(ny_noaa)`__. The __proportion of missing value__ in each column is shown in the table above. As each weather station may collect only a subset of these variables, the resulting dataset contains extensive missing data. The __Maximum temperature and Minimum temperature__ hold the highest proportion of missing value, which are almost half of the observations in each column. Meanwhile, __Precipitation__ has a lower proportion of missing values compared to others. The big __difference of proportion__ (0.056 vs. 0.437) is a big issue. 

### 3.2

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?

```{r clean, creat new observations}
# clean the data
nynoaa_clean = ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(month, month = as.numeric(month)) %>% # separate dat
  mutate(prcp = prcp / 10,
         tmax = as.numeric(tmax) / 10,
         tmin = as.numeric(tmin) / 10) # adjust units 
# function for mode
getmode = function(x) {
  ux = unique(x)
  ux[which.max(tabulate(match(x, ux)))]}

getmode(nynoaa_clean$snow)
```

The most commonly observed values in snowfall is __0__. Because snowfall happens in Winter days, most of the time in a year will have 0 snowfall.

### 3.3

Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r twoplot, fig.width = 10,fig.height = 15}
# arrange the data
tmax_meantb = nynoaa_clean %>% 
  filter(month == 01 | month == 07) %>% 
  group_by(id, year, month) %>% 
  summarize(mean_tmax = round(mean(tmax, na.rm = TRUE), digits = 3))
# creat the labeller
month_lab = c('1' = "January", '7' = "July")
# make the plot
tmax_meantb %>% 
  ggplot(aes(x = year, y = mean_tmax, group = id, color = id)) + 
  geom_point(size = 0.5) + 
  geom_line(size = 0.3, alpha = 0.6) + 
  facet_grid(~month, labeller = labeller(month = month_lab)) +
  theme_bw() +
  theme(axis.text.x = element_text(size = 5, angle = 90), legend.position = "hide", plot.title = element_text(hjust = 0.5)) +
  labs(title = "1981-2010 Average Max Temperature in January and July in Each Station ",
       x = "Year",
       y = "Average Max Temperature (˚C)",
       caption = "Data from NOAA National Climatic Data Center") 
```

In the plot, we can see that the most of the average max temperature in January are __from -10˚C to 10˚C__ in each station across years and in July are __from 20˚C to 35˚C__. However, in both months, the __changes were flutant__ from 1981-2010. We can see the lines goes up and down. From __July 2009 to July 2010__, there is a siginificant growth, which may reflect the abnormal changes in climate. As mean is affected by extreme value, more analyses are needed. 

There are some outliners stations in both months. For example, there are one in __January 1982__ and one in __July 1988__.

### 3.4

Make a two-panel plot showing 

(i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option)

(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r two panel plot}
# make the plot 1
p1 =  nynoaa_clean %>% 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex() + 
  theme_bw() +
  theme(legend.position = "bottom", plot.title = element_text(size = 6, hjust = 0.5), legend.text = element_text(size = 3)) +
  labs(
    title = "1981-2010 The Comparison of Max Temperature and Min Temperature in Each Station",
    x = "Maximum Temperature(˚C)",
    y = "Minimum Temperature(˚C)")
# make the plot 2
p2 = nynoaa_clean %>% 
  group_by(year) %>% 
  mutate(range = snow > 0 & snow < 100) %>% 
  filter(range == TRUE, !is.na(range)) %>% 
  select(id, year, snow) %>% 
  ggplot(aes(x = snow, fill = year)) + 
  geom_density(alpha = 0.3) +
  theme_bw() +
  theme(legend.position = "hide", plot.title = element_text(hjust = 0.5, size = 6)) +
  labs(
    title = "The Density Distribution of Snowfall in Each Year (> 0mm & < 100mm)",
    x = "Snow Fall (mm)",
    y = "Density",
    caption = "Data from NOAA National Climatic Data Center")
# combine the two plot
twopanel = p1 + p2
twopanel
```

From the plots above, we can see that in most of the stations, max and min temperature are between -10˚C to 25˚C. There are small amount of snow fall observed in these stations as the density of snow fall between 0 to 25mm are high.

