---
title: "p8105_hw3_zf2212"
author: "Catherine"
date: "10/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
library(tidyverse)
```

## Problem 1 BRFSS

First, do some data cleaning:

```{r import and tidy data}
# import
library(p8105.datasets)
data("brfss_smart2010")

brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% # use appropriate variable names
  filter(topic == "Overall Health") %>% # filter topic 
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor"), !is.na("response")) %>% 
  # filter response from excellent to poor
  mutate(response, as.factor(response)) %>% 
  arrange(response) %>%  # mutate response as factor and arrange order
  select()
  

```

Using this dataset, do or answer the following (commenting on the results of each):
In 2002, which states were observed at 7 locations?
Make a “spaghetti plot” that shows the number of locations in each state from 2002 to 2010.
Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.